GVN-Hoist: Hoisting Computations from Branches

### 2016-08-08 hoist.tex
profitability

\subsection{Checking for legality}
Finding a hoisting point i.e. a common dominator of the set of instructions. If
such a dominator is found then checking if all the use-operands of the set of
instructions are available or not. In some cases if the operands are not
available, it is possible to reinstantiate (remateralize) the use-operands, thus
making the passing the legality check. Checking that the side-effects of the
computations does not intersect with any side-effect between the instructions to
be hoisted and their hoisting point. For memory operations like loads, stores,
calls etc., it is also required to check that all the paths from the hoisting
point to the end of the function should have the exact memory operation
available, in order to guarantee correctness.

\subsubsection{Hoisting scalars}
Scalars are the easiest to hoist because we do not have to analyze them for
aliasing memory references. As long as all the operands are available, the
scalar can be hoisted. Care must be taken in case of hoisting scalars too far,
as that may increase register pressure and result in spills. For example
hoisting a scalar past a call (Section-\ref{cost:across-calls}). In our current
implementation we hoist scalars past a call only when optimizing for code-side
(-Os). Another way to mitigate this problem is be to reinstantiate
(rematerialize) the computation after a call (may be as a different pass).

\subsubsection{Hoisting loads}
In order to value number stores, we value number (hash) the address from where
the value is to be loaded. Then we need to compute the hoisting point similar to
the hoisting of scalars. At the hoisting point the availability of operands to the
load (an address) is checked. If it is not available we try to compute the
addresss if that is possible.  Along the path, from current position of the load
instruction through the hoisting point, we find if there are aliasing stores, we
find if any of the basic blocks are landing pads or targets of indirect
branches. In those cases we discard the candidate. Since the load increases the
register pressure by one (unless the operand to load becomes dead at the load),
we also evaluate if it is beneficial to hoist the load.

\subsubsection{Hoisting stores}
In order to value number stores, we value number (hash) the address as well as
the value to be stored at that address. For the legality, we need to check the
dependency requirements similar to the hoisting of loads. We need to check that
the operands of the store instruction are available at the hoisting point, that
there are no aliasing loads or store along the path from the current position
through the hoisting point and, none of the basic blocks along the path have
their address taken. Once all the legality checks are satisfied, it is a good
candidate to be hoisted.

\subsubsection{Hoisting calls}
llvm has built in facility to value-number calls. If there are calls in sibling
branches with identical value numbers, they can be hoisted. For legality we have
to check that the loads/stores in the call does not create hazard with any
loads/stores (in the caller) in the path from callsite through the hoisting
point, just like we did for loads and stores.

@@

>>>>

\subsection{Checking for legality}
Finding a hoisting point i.e., a common dominator of the set of instructions. If
such a dominator is found then checking if all the use-operands of the set of
instructions are available or not. In some cases if the operands are not
available, it is possible to reinstantiate (remateralize) the use-operands, thus
making the passing the legality check. Checking that the side-effects of the
computations does not intersect with any side-effect between the instructions to
be hoisted and their hoisting point. For memory operations like loads, stores,
calls etc., it is also required to check that all the paths from the hoisting
point to the end of the function should have the exact memory operation
available, in order to guarantee correctness.

\subsubsection{Hoisting scalars}
Scalars are the easiest to hoist because we do not have to analyze them for
aliasing memory references. As long as all the operands are available, the
scalar can be hoisted.

\subsubsection{Hoisting loads}
At the hoisting point the availability of operands to the load (an address) is
checked. If it is not available we try to compute the addresss if that is
possible.  Along the path, from current position of the load instruction through
the hoisting point, we find if there are aliasing stores, we find if any of the
basic blocks are landing pads or targets of indirect branches. In those cases we
discard the candidate.

\subsubsection{Hoisting stores}
We need to check the dependency requirements similar to the hoisting of
loads. We need to check that the operands of the store instruction are available
at the hoisting point, that there are no aliasing loads or store along the path
from the current position through the hoisting point and, none of the basic
blocks along the path have their address taken.

\subsubsection{Hoisting calls}
We have to check that the loads/stores in the call does not create hazard with
any loads/stores (in the caller) in the path from callsite through the hoisting
point, just like we did for loads and stores.

@@

\subsubsection{Hoisting scalars}
Care must be taken in case of hoisting scalars too far, as that may increase
register pressure and result in spills. For example hoisting a scalar past a
call (Section-\ref{cost:across-calls}). In our current implementation we hoist
scalars past a call only when optimizing for code-side (-Os). Another way to
mitigate this problem is be to reinstantiate (rematerialize) the computation
after a call (may be as a different pass).

\subsubsection{Hoisting loads}
A load instruction introduces a register where the value loaded will be kept,
the register pressure increases by one (unless the operand to load becomes dead
at the load). On the other hand, loading a value early will reduce the stall
during execution should the value is not in the cache. We generally prefer to
hoist load except the hoisting point is too far (this distance is computed by
looking at the experimental results of representative benchmarks see
Section-\ref{sec:experimental-results}).

\subsubsection{Hoisting stores}
Since stores do not increase the live-range of any registers, and in some cases
it ends the liveness of registers, we hoist all the stores.

\subsubsection{Hoisting calls}
Currently we hoist all the calls.
